### 2022-12-21 C3.tex
Add files via upload

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2013 International Union of Crystallography
% Version 1.6 (28 March 2013)
%------------------------------------------------------------------------------

\documentclass[]{iucr}              % DO NOT DELETE THIS LINE
%\documentclass[]{iucr}              % DO NOT DELETE THIS LINE
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
%\usepackage{bm}
\usepackage{graphicx}
\usepackage{tabularx}
%\usepackage{calligra}
\usepackage{array}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}
\def\mathbi#1{\textbf{\em #1}}
\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\hyphenation{Niggli}
\def\mathbi#1{\textbf{\em #1}}
%\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\usepackage{color}
\usepackage{ulem}
\usepackage{url}
\usepackage{yfonts}
%\usepackage{xr-hyper}
%\usepackage[draft]{hyperref}
%\usepackage{bibentry}
\newcommand{\SVI}[0]{$\bf{S^{6}}$}
\newcommand{\GVI}[0]{$\bf{G^{6}}$}
\newcommand{\CIII}[0]{$\bf{C^{3}}$}
\newcommand{\DVII}[0]{$\bf{D^{7}}$}
\newcommand{\VVII}[0]{$\bf{V^{7}}$}

\newcommand{\vdotv}[2]{${{\bf #1 \cdot #2}}$}
\newcommand{\Imaginary}[0]{$ \textfrak{I} $}
\newcommand{\Real}[0]{$ \textfrak{R} $}
\newcommand{\Exchange}[0]{$\textfrak{X}$}

\newcommand{\nounderline}[3]{\!\!\!\!\!\!\!\!\!#1,&\!\!\!\!\!\!\!\!\!#2,&\!\!\!\!\!\!\!\!\!#3}
\newcommand{\underlineab}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#1),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#2),&\!\!\!\!\!\!\!\!\!#3}
\newcommand{\underlineac}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#1),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!#2,&\!\!\!\!\!\!\!\!\!\Exchange{}(#3)}
\newcommand{\underlinebc}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!#1,&\!\!\!\!\!\!\!\!\!\Exchange{}(#2),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#3)}

\newcommand{\scalar}[1]{$#1$}

\newcommand{\scalarsub}[2]{$#1_#2$}

%-------------------------------------------------------------------------
% Information about journal to which submitted
%-------------------------------------------------------------------------
\journalcode{A}              % Indicate the journal to which submitted
%   A - Acta Crystallographica Section A
%   B - Acta Crystallographica Section B
%   C - Acta Crystallographica Section C
%   D - Acta Crystallographica Section D
%   E - Acta Crystallographica Section E
%   F - Acta Crystallographica Section F
%   J - Journal of Applied Crystallography
%   M - IUCrJ
%   S - Journal of Synchrotron Radiation
\makeatletter
\font\dummyft@=dummy \relax
\makeatother

\begin{document}                  % DO NOT DELETE THIS LINE

	%-------------------------------------------------------------------------
	% The introductory (header) part of the paper
	%-------------------------------------------------------------------------

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).

	% Authors' names and addresses. Use \cauthor for the main (contact) author.
	% Use \author for all other authors. Use \aff for authors' affiliations.
	% Use lower-case letters in square brackets to link authors to their
	% affiliations; if there is only one affiliation address, remove the [a].

	% Use \vita if required to give biographical details (for authors of
	% invited review papers only). Uncomment it.

	% lca IUCr id IUCr6401
	%\vita{Author's biography}

	% Keywords (required for Journal of Synchrotron Radiation only)
	% Use the \keyword macro for each word or phrase, e.g. 
	% \keyword{X-ray diffraction}\keyword{muscle}

	% PDB and NDB reference codes for structures referenced in the article and
	% deposited with the Protein Data Bank and Nucleic Acids Database (Acta
	% Crystallographica Section D). Repeat for each separate structure e.g
	% \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

	%\PDBref[optional name]{refcode}
	%\NDBref[optional name]{refcode}

	%-------------------------------------------------------------------------
	% The introductory (header) part of the paper
	%-------------------------------------------------------------------------

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).
	{\LARGE \emph{\today}} \\
	\title{Delone lattice studies in \CIII, the space of three complex variables}
	%\title{Note on the transformation of three-space basis vectors to  corresponding matrix for Delaunay scalars}
	\shorttitle{properties of C3}

	% Authors' names and addresses. Use \cauthor for the main (contact) author.
	% Use \author for all other authors. Use \aff for authors' affiliations.
	% Use lower-case letters in square brackets to link authors to their
	% affiliations; if there is only one affiliation address, remove the [a].

	\cauthor[a]{Lawrence C.}{Andrews}{lawrence.andrews@ronininstitute.org}{}
	\author[b]{Herbert J.}{Bernstein}

	\aff[a]{Ronin Institute, 9515 NE 137th St, Kirkland, WA, 98034-1820 \country{USA}}
	\aff[b]{Ronin Institute, c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}}

	% Use \shortauthor to indicate an abbreviated author list for use in
	% running heads (you will need to uncomment it).

	\shortauthor{Andrews and Bernstein}

	% Use \vita if required to give biographical details (for authors of
	% invited review papers only). Uncomment it.

	% lca IUCr id IUCr6401
	%\vita{Author's biography}

	% Keywords (required for Journal of Synchrotron Radiation only)
	% Use the \keyword macro for each word or phrase, e.g. 
	% \keyword{X-ray diffraction}\keyword{muscle}

	\keyword{Delaunay}
	\keyword{Delone}
	\keyword{Selling}
	\keyword{\CIII}

	% PDB and NDB reference codes for structures referenced in the article and
	% deposited with the Protein Data Bank and Nucleic Acids Database (Acta
	% Crystallographica Section D). Repeat for each separate structure e.g
	% \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

	%\PDBref[optional name]{refcode}
	%\NDBref[optional name]{refcode}

	\maketitle                        % DO NOT DELETE THIS LINE

	\begin{synopsis}
		Selling reduction and Delone reduction are
		considered in a space of complex variables.
	\end{synopsis}
	\newcommand{\si}[0]{$s_1$}
	\newcommand{\sii}[0]{$s_2$}
	\newcommand{\siii}[0]{$s_3$}
	\newcommand{\siv}[0]{$s_4$}
	\newcommand{\sv}[0]{$s_5$}
	\newcommand{\svi}[0]{$s_6$}
	\newcommand{\Svec} [0] {\{\si, \sii, \siii, \siv, \sv, \svi \}}
	\newcommand{\SvecA} [0] {\{-\si, -\si+\sii, \si+\siii, \si+\sv, \si+\siv, \si+\svi \}}

	\newcommand{\OPES}[0]{$E^3toS^6$}
	\newcommand{\OPESS}[0]{$$E^3toS^6$$}
	\newcommand{\MSVI}[0]{$M_{S^{6}}$}
	\newcommand{\MEIII}[0]{$M_{E^{3}}$}
	\newcommand{\Plus}[0]{$\textfrak{P}$}	
	\newcommand{\Minus}[0]{$\textfrak{M}$}

	\newcommand{\ci}[0]{$c_1$}
	\newcommand{\cii}[0]{$c_2$}
	\newcommand{\ciii}[0]{$c_3$}

	\begin{abstract}

		The Delone scalars are studied in \CIII, the space of three complex variables.

		{\bf Note:}  In his later publications, Boris Delaunay used the Russian version of his surname, Delone.\\

	\end{abstract}

	%-------------------------------------------------------------------------
	% The main body of the paper
	%-------------------------------------------------------------------------
	% Now enter the text of the document in multiple \section's, \subsection's
	% and \subsubsection's as required.

	% Appendices appear after the main body of the text. They are prefixed by
	% a single \appendix declaration, and are then structured just like the
	% body text.

	\section{Introduction}

	The scalars used by \citeasnoun{Delaunay1932} in his formulation of Selling reduction \cite{Selling1874}
	are (in the conventional order) $b \cdot c$, $a \cdot c$, $a \cdot b$, $a \cdot d$, 
	$b \cdot d$, $c \cdot d$, where $d = -a-b-c$. (As a mnemonic device, observe that the first three terms use
	$\alpha$, $\beta$, and $\gamma$, in that order, and the following terms use $a$, $b$, $c$, in that order.)

	\citeasnoun{andrews2019b} chose to represent the scalars in the space \SVI{},
	\Svec{} (defined in the order above), as a way to create a metric space
	for the measurement of the distance between lattices. They also consider the representation of this space as the
	space of three complex dimensions, \CIII{} or {$\{c_1$}, {$c_2$}, {$c_3\}$}. 	
	In \CIII{}, in terms of the Delone scalars, a vector is \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}, where the real and imaginary parts
	of each are the ``opposite'' scalars according to the definition of \citeasnoun{Delaunay1932} (see \citeasnoun{andrews2019a}).
	As a memonic device, note that the complex components involve ($\alpha,a$), ($\beta, b$), and ($\gamma,c$).

		\citeasnoun{andrews2019b} considered the matrix representations of the reflections in \CIII{} (and in \SVI{}). This paper will describe the boundary transformations at the edges of the fundamental
	unit of \CIII{}. In \SVI{}, the fundamental unit is the all negative orthant, which contains only and
	all of the reduced cells. In \SVI{} and \CIII{}, the boundaries are found where any scalar (or correspondingly the real 
	or imaginary part) is equal to zero. The rationale for this work was that it might lend insights into the
	topology of the space of lattices.

	\section{Notation}

	Complex numbers will be represented in Cartesian format $(x,y)$, where $x$ is the real part and
	$y$ is the imaginary part.

	We will represent a vector in \CIII{} by \{$(x_1,y_1)$, $(x_2,y_2)$, $(x_3, y_3)$\} as an alternative to  \{(\si,\siv ), (\sii,\sv),(\sii,\svi)\}.

	Next we define the operators in \CIII{} that will be used in the matrix descriptions of 
	the transformations at the boundaries of the fundamental unit.\\

	\begin{table}	
		\begin{tabular}{|l|l|l|l|}
			\hline
			Operator&Usage&Result&Name \\ \hline
			$\Minus{}_r $&  $\Minus{}_r(c_j)$ & $(-x_j, -x_j+y_j)$ & Minus real \\ \hline
			$\Minus{}_i $ & $\Minus{}_i(c_j)$ & $(x_j-y_j,-y_j) $ & Minus imag\\ \hline
			$\textfrak{P}_r$ & $ \textfrak{P}_r(c_j)$ & $(x_j, x_j)$ & Plus real\\ \hline
			$\textfrak{P}_i$ &  $\textfrak{P}_i(c_j)$  & $(y_j, y_j)$ & Plus imag\\ \hline
			$\Re$ &  $\Re(c_j)$ & $x_j$ & Real \\ \hline
			$\Im$  &$ \Im(c_j)$ & $y_j$ & Imaginary \\ \hline

		\end{tabular}

	\end{table}

	\section{Matrices of boundary transformations}

	For the boundary at $s_1$: (the real component of $c_1$).

	$\begin{bmatrix}
	\Minus{}_r & 0 & 0 \\
	\Plus{}_r &  \textrm{i}\Re{} & \Re{} \\
	\textfrak{P}_r & \textrm{i}\Im{} & \Im{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
	\Minus{}_r & 0 & 0 \\
	\Plus{}_r &\textrm{i}\Im{} & \Im{} \\
	\Plus{}_r & \textrm{i}\Re{} & \Re{} \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
	\Minus{}_r & 0 & 0 \\
	\Plus{}_r &  \Re{} & \textrm{i}\Re{} \\
	\Plus{}_r & \Im{} &\textrm{i}\Im{}
	\end{bmatrix}	$
	%
	$\begin{bmatrix}
	\Minus{}_r & 0 & 0 \\
	\Plus{}_r & \Im{} &\textrm{i}\Im{} \\
	\Plus{}_r &  \Re{} &\textrm{i}\Re{} \\
	\end{bmatrix}$ \\

	For the boundary at $s_4$: (the imaginary component of $c_1$).

	$\begin{bmatrix}
	\Minus{}_i & 0 & 0 \\
	\Plus{}_i &  \textrm{i}\Re{} & \Re{} \\
	\Plus{}_i & \textrm{i}\Im{} & \Im{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
	\Minus{}_i & 0 & 0 \\
	\Plus{}_i & \textrm{i}\Im{} & \Im{} \\
	\Plus{}_i &  \textrm{i}\Re{} & \Re{} \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
	\Minus{}_i & 0 & 0 \\
	\Plus{}_i &  \Re{} & \textrm{i}\Re{} \\
	\Plus{}_i & \Im{} & \textrm{i}\Im{}
	\end{bmatrix}	$
	%
	$\begin{bmatrix}
	\Minus{}_i & 0 & 0 \\
	\Plus{}_i & \Im{} & \textrm{i}\Im{} \\
	\Plus{}_i &  \Re{}& \textrm{i}\Re{} \\
	\end{bmatrix}$ \\

	For the boundary at \sii{} (the real component of \cii{}):

	$\begin{bmatrix}

	\textrm{i}\Re{} & \Plus{}_r & \Re{} \\
	0 & \Minus{}_r & 0 \\
	\textrm{i}\Im{} & \Plus{}_r & \Im{}
\end{bmatrix}$
%
$\begin{bmatrix}
	\textrm{i}\Im{} & \Plus{}_r &  \Im{} \\
	0 & \Minus{}_r & 0 \\
	\textrm{i}\Re{} & \Plus{}_r & \Re{} \\
\end{bmatrix}$ 
%
$\begin{bmatrix}
	\Re{} &\Plus{}_r &   \textrm{i}\Re{} \\
	0 & \Minus{}_r &  0 \\
	\Im{} &\Plus{}_r & \textrm{i}\Im{}
\end{bmatrix}	$ 
%
$\begin{bmatrix}
	\Im{} & \Plus{}_r & \textrm{i}\Im{} \\
	0 &\Minus{}_r &  0 \\
	\Re{} & \Plus{}_r &  \textrm{i}\Re{} \\
\end{bmatrix}$ \\

For the boundary at $s_5$: (the imaginary component of \cii{}).

$\begin{bmatrix}
\textrm{i}\Re{} & \Plus{}_i & \Re{} \\
0 & \Minus{}_i & 0 \\
\textrm{i}\Im{} & \Plus{}_i & \Im{}
\end{bmatrix}$
%
$\begin{bmatrix}
\textrm{i}\Im{} & \Plus{}_i &  \Im{} \\
0 & \Minus{}_i & 0 \\
\textrm{i}\Re{} & \Plus{}_ri& \Re{} \\
\end{bmatrix}$ 
%
$\begin{bmatrix}
\Re{} &\Plus{}_i &   \textrm{i}\Re{} \\
0 & \Minus{}_i &  0 \\
\Im{} &\Plus{}_i & \textrm{i}\Im{}
\end{bmatrix}	$ 
%
$\begin{bmatrix}
\Im{} & \Plus{}_i & \textrm{i}\Im{} \\
0 &\Minus{}_i &  0 \\
\Re{} & \Plus{}_i &  \textrm{i}\Re{} \\
\end{bmatrix}$ \\

For the boundary at \siii{} (the real component of \ciii{}):

$\begin{bmatrix}

\textrm{i}\Re{} & \Re{} & \Plus{}_r \\
\textrm{i}\Im{} & \Im{}& \Plus{}_r  \\
0 & 0& \Minus{}_r  \\
\end{bmatrix}$
%
$\begin{bmatrix}
\textrm{i}\Im{} &  \Im{} & \Plus{}_r \\
\textrm{i}\Re{} & \Re{} & \Plus{}_r \\
0 & 0 & \Minus{}_r \\
\end{bmatrix}$ 
%
$\begin{bmatrix}
\Re{} &   \textrm{i}\Re{} &\Plus{}_r \\
\Im{} & \textrm{i}\Im{}&\Plus{}_r \\
0  &  0 & \Minus{}_r \\
\end{bmatrix}	$ 
%
$\begin{bmatrix}
\Im{} & \textrm{i}\Im{} & \Plus{}_r \\
\Re{} &  \textrm{i}\Re{} & \Plus{}_r \\
0 &  0 &\Minus{}_r \\
\end{bmatrix}$ \\

For the boundary at \svi{} (the imaginary component of \ciii{}):

$\begin{bmatrix}

\textrm{i}\Re{} & \Re{} & \Plus{}_i \\
\textrm{i}\Im{} & \Im{}& \Plus{}_i \\
0 & 0 & \Minus{}_i 
\end{bmatrix}$
%
$\begin{bmatrix}
\textrm{i}\Im{} &  \Im{} & \Plus{}_i \\
\textrm{i}\Re{} & \Re{} & \Plus{}_i \\
0 & 0 & \Minus{}_i \\
\end{bmatrix}$ 
%
$\begin{bmatrix}
\Re{} &   \textrm{i}\Re{}  &\Plus{}_i \\
\Im{} & \textrm{i}\Im{} &\Plus{}_i \\
0 &  0 & \Minus{}_i \\
\end{bmatrix}	$ 
%
$\begin{bmatrix}
\Im{} & \textrm{i}\Im{} & \Plus{}_i \\
\Re{} &  \textrm{i}\Re{} & \Plus{}_i \\
0 &  0 &\Minus{}_i 
\end{bmatrix}$ \\

\section{Basics}

The standard representation of the identity operation is

	$\bf{c'}$  = 
	$\begin{bmatrix}
	$1$ & $0$ &  $0$ \\
	$0$ & $1$ & $0$ \\
	$0$ & $0$ & $1$ \\
	\end{bmatrix}$ 
	$\bf{c}$ \\

The identity in \CIII{} can also be written:

	$\bf{c'}$  = 
	$\begin{bmatrix}
	$1$ & $0$ &  $0$ \\
	$0$ & \Re{} +\textrm{i}\Im{} & $0$ \\
	$0$ & $0$ & \Re{} +\textrm{i}\Im{} \\
	\end{bmatrix}$ 
	$\bf{c}$
\\

\citeasnoun{Delaunay1932} does not consider the boundary transformations 
in detail. However, he uses them to define the process of Selling reduction. 
For example in \SVI{}, he lists the following as one of the possible results 
for a transformation on \si{}.
\SvecA{}
The third matrix above (the third boundary transform for $s_1$) implements this operation and interchanges the
real part of \ciii{} and the imaginary part of \cii{}.
~\\
	$\begin{bmatrix}
\Minus{}_r & 0 & 0 \\
\Plus{}_r &  \Re{} & \textrm{i}\Re{} \\
\Plus{}_r & \Im{} &\textrm{i}\Im{}
\end{bmatrix}	$
\\

In terms of \CIII, Delone's alternate transformation for the \si{} boundary would
exchange the real of \cii{} with the imaginary part of \ciii{}. That is the fourth
matrix in the list for \si{}. The other two transformations for \si{} can 
be generated from the two we have just displayed by the "exchange operation" \cite{andrews2019b} applied to the second and third \CIII{} elements. Delone 
did not describe the latter two transformation, perhaps because even a
single transformation was adequate to implement reduction. He already listed
two.

\section{Summary}

The considerable regularity of the transformation matrices (above) highlights one
of the important aspects of Selling reduction as used by \citeasnoun{Delaunay1932}
when compared to Niggli reduction \cite{Niggli1928}. Where all the 
boundaries of the fundamental orthant in \CIII ~and \SVI{} are basically the same (and indeed are 
related by reflections), the boundaries formed in Niggli reduction
are of multiple types, and the fundamental unit of the space
representing Niggli reduced cells (\GVI{}, see \citeasnoun{Andrews2014}) is non-convex.

The matrices display one of the properties of \CIII{} that make it a
useful representation. For the scalar being transformed, it and its 
opposite scalar create a unique row in each matrix. Those rows contain
only the minus operator and zeros, highlighting the unique relationship
of that pair of scalars.

Several aspects of C3 are obvious from inspecting the matrices above. 
First, we see that each boundary has four possible 
transformations that
can be applied. Because the each of the transformations at boundaries are 
self-inverse, they are the same transformations that would 
be used in the process of cell (lattice) reduction. 
\citeasnoun{Delaunay1932} and  \citeasnoun{Delone1975} give 
only two choices. Presumably for simplicity they did not list
the transformations that used the "exchange" operator 
(see \cite{andrews2019b}).

\section{Availability of code}

The $C^{++}$ ~code for \CIII{} is available in github.com, in
\url{https://github.com/duck10/LatticeRepLib.git}.

%\appendix

%\section{blah blah blah -- Supplementary Material}
\ack{{\bf Acknowledgements}}

Careful copy-editing and corrections by Frances C. Bernstein are 
gratefully acknowledged.	Our thanks to Jean Jakoncic and Alexei Soares for 
helpful conversations and access to data and facilities at 
Brookhaven National Laboratory.

\ack{{\bf Funding information}}

Funding for this research was provided in part by:  
US Department of Energy Offices of Biological and 
Environmental Research and of Basic Energy Sciences 
(grant No. DE-AC02-98CH10886; grant No. E-SC0012704); 
U.S. National Institutes of Health (grant No. P41RR012408; 
grant No. P41GM103473; grant No. P41GM111244; 
grant No. R01GM117126,
grant No. 1R21GM129570); Dectris, Ltd.

\bibliography{reduced}

\bibliographystyle{iucr}

%-------------------------------------------------------------------------
% TABLES AND FIGURES SHOULD BE INSERTED AFTER THE MAIN BODY OF THE TEXT
%-------------------------------------------------------------------------

% Simple tables should use the tabular environment according to this
% model

% Postscript figures can be included with multiple figure blocks

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%