### 2023-01-06 ProgramSummary/ProgramSummary..tex
continuing work on command line program writeup

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2013 International Union of Crystallography
% Version 1.6 (28 March 2013)
%------------------------------------------------------------------------------

%xxxxxx just for testing github % 2
%xxxxxx just for testing github % 3

\documentclass[preprint]{iucr}              % DO NOT DELETE THIS LINE
%\documentclass[]{iucr}              % DO NOT DELETE THIS LINE
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
%\usepackage{xtabular]
%\usepackage{bm}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{calligra}
\usepackage{array}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}
\def\mathbi#1{\textbf{\em #1}}
\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\hyphenation{Niggli}
\def\mathbi#1{\textbf{\em #1}}
%\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\usepackage{color}
\usepackage{ulem}
\usepackage{url}
\usepackage{yfonts}
%\usepackage{xr-hyper}
%\usepackage[draft]{hyperref}
%\usepackage{bibentry}
\newcommand{\BIV}[0]{$\bf{B^{4}}$}
\newcommand{\SVI}[0]{$\bf{S^{6}}$}
\newcommand{\GVI}[0]{$\bf{G^{6}}$}
\newcommand{\CIII}[0]{$\bf{C^{3}}$}
\newcommand{\DVII}[0]{$\bf{D^{7}}$}
\newcommand{\DXIII}[0]{$\bf{D^{13}}$}
\newcommand{\VVII}[0]{$\bf{V^{7}}$}

\newcommand{\vdotv}[2]{${{\bf #1 \cdot #2}}$}
\newcommand{\Imaginary}[0]{$ \textfrak{I} $}
\newcommand{\Real}[0]{$ \textfrak{R} $}
\newcommand{\Exchange}[0]{$\textfrak{X}$}

\newcommand{\nounderline}[3]{\!\!\!\!\!\!\!\!\!#1,&\!\!\!\!\!\!\!\!\!#2,&\!\!\!\!\!\!\!\!\!#3}
\newcommand{\underlineab}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#1),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#2),&\!\!\!\!\!\!\!\!\!#3}
\newcommand{\underlineac}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#1),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!#2,&\!\!\!\!\!\!\!\!\!\Exchange{}(#3)}
\newcommand{\underlinebc}[3]{\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!#1,&\!\!\!\!\!\!\!\!\!\Exchange{}(#2),&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\Exchange{}(#3)}

\newcommand{\scalar}[1]{$#1$}

\newcommand{\scalarsub}[2]{$#1_#2$}

%-------------------------------------------------------------------------
% Information about journal to which submitted
%-------------------------------------------------------------------------
%\journalcode{A}              % Indicate the journal to which submitted
%   A - Acta Crystallographica Section A
%   B - Acta Crystallographica Section B
%   C - Acta Crystallographica Section C
%   D - Acta Crystallographica Section D
%   E - Acta Crystallographica Section E
%   F - Acta Crystallographica Section F
%   J - Journal of Applied Crystallography
%   M - IUCrJ
%   S - Journal of Synchrotron Radiation
\makeatletter
\font\dummyft@=dummy \relax
\makeatother

\begin{document}                  % DO NOT DELETE THIS LINE

	%-------------------------------------------------------------------------
	% The introductory (header) part of the paper
	%-------------------------------------------------------------------------

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).

	% Authors' names and addresses. Use \cauthor for the main (contact) author.
	% Use \author for all other authors. Use \aff for authors' affiliations.
	% Use lower-case letters in square brackets to link authors to their
	% affiliations; if there is only one affiliation address, remove the [a].

	% Use \vita if required to give biographical details (for authors of
	% invited review papers only). Uncomment it.

	% lca IUCr id IUCr6401
	%\vita{Author's biography}

	% Keywords (required for Journal of Synchrotron Radiation only)
	% Use the \keyword macro for each word or phrase, e.g. 
	% \keyword{X-ray diffraction}\keyword{muscle}

	% PDB and NDB reference codes for structures referenced in the article and
	% deposited with the Protein Data Bank and Nucleic Acids Database (Acta
	% Crystallographica Section D). Repeat for each separate structure e.g
	% \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

	%\PDBref[optional name]{refcode}
	%\NDBref[optional name]{refcode}

	%-------------------------------------------------------------------------
	% The introductory (header) part of the paper
	%-------------------------------------------------------------------------

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).
	{\LARGE \emph{\today}} \\
	\title{Command line programs from LatticeRepLib}
	\shorttitle{Command line programs}

	% Authors' names and addresses. Use \cauthor for the main (contact) author.
	% Use \author for all other authors. Use \aff for authors' affiliations.
	% Use lower-case letters in square brackets to link authors to their
	% affiliations; if there is only one affiliation address, remove the [a].

	\cauthor[a]{Lawrence C.}{Andrews}{lawrence.andrews@ronininstitute.org}{}
	\author[b]{Herbert J.}{Bernstein}

	\aff[a]{Ronin Institute, 9515 NE 137th St, Kirkland, WA, 98034-1820 \country{USA}}
	\aff[b]{Ronin Institute, c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}}

	% Use \shortauthor to indicate an abbreviated author list for use in
	% running heads (you will need to uncomment it).

	\shortauthor{Andrews and Bernstein}

	% Use \vita if required to give biographical details (for authors of
	% invited review papers only). Uncomment it.

	% lca IUCr id IUCr6401
	%\vita{Author's biography}

	% Keywords (required for Journal of Synchrotron Radiation only)
	% Use the \keyword macro for each word or phrase, e.g. 
	% \keyword{X-ray diffraction}\keyword{muscle}

	%	\keyword{Delaunay}
	%	\keyword{Delone}
	%	\keyword{Selling}
	%	\keyword{\CIII}

	% PDB and NDB reference codes for structures referenced in the article and
	% deposited with the Protein Data Bank and Nucleic Acids Database (Acta
	% Crystallographica Section D). Repeat for each separate structure e.g
	% \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

	%\PDBref[optional name]{refcode}
	%\NDBref[optional name]{refcode}
	\maketitle                        % DO NOT DELETE THIS LINE

	\begin{synopsis}
		Selling reduction and Delone reduction are
		considered in a space of complex variables.
	\end{synopsis}
	\newcommand{\si}[0]{$s_1$}
	\newcommand{\sii}[0]{$s_2$}
	\newcommand{\siii}[0]{$s_3$}
	\newcommand{\siv}[0]{$s_4$}
	\newcommand{\sv}[0]{$s_5$}
	\newcommand{\svi}[0]{$s_6$}
	\newcommand{\Svec} [0] {\{\si, \sii, \siii, \siv, \sv, \svi \}}
	\newcommand{\SvecA} [0] {\{-\si, -\si+\sii, \si+\siii, \si+\sv, \si+\siv, \si+\svi \}}

	\newcommand{\OPES}[0]{$E^3toS^6$}
	\newcommand{\OPESS}[0]{$$E^3toS^6$$}
	\newcommand{\MSVI}[0]{$M_{S^{6}}$}
	\newcommand{\MEIII}[0]{$M_{E^{3}}$}
	\newcommand{\Plus}[0]{$\textfrak{P}$}	
	\newcommand{\Minus}[0]{$\textfrak{M}$}

	\newcommand{\ci}[0]{$c_1$}
	\newcommand{\cii}[0]{$c_2$}
	\newcommand{\ciii}[0]{$c_3$}

	\section{Introduction}

	In the study of various spaces for representing lattices, a
	number of software tools have been prepared. Some of those
	are available as simple commandline programs. Most of them
	have flexible input and output their processed results in 
	a form that another can use that for processing. A few
	are terminal programs that produce analysis. A small
	number take no input and generate files for other uses.

	% Appendices appear after the main body of the text. They are prefixed by
	% a single \appendix declaration, and are then structured just like the
	% body text.

	\section{Data Inputs:}

	\begin{table}
		\caption{	All these are case-insensitive. If a particular
			input lattice is invalid, it is rejected with a message.}
		\begin{tabular}{l p{.2\textwidth} l l l}
			\toprule

			\begin{tabular}{l l l l}
				\underline{\textbf{Vector Input:}}&
				\parbox[t]{0.6\textwidth}{\textbf{g} 
					(or \textbf{v} or \textbf{g6}) for \GVI{} vectors\\
					\textbf{d} (or \textbf{d7}) for \DVII{} vectors\\
					\textbf{s} (or s6) for \SVI{}, Delone/Selling scalars\\
					\textbf{C3} for \CIII{} input (without parentheses or commas, \\
					~~~“C” would be interpreted as a C-centered unit cell)}
			\end{tabular}	\\[3pt]

			\underline{\textbf{RANDOM:}} Random (valid) unit cell generated\\

			\textbf{Crystal Lattice Input:} 
			``\textbf{A}'', ``\textbf{B}'', ``\textbf{C}'', ``\textbf{P}'', ``\textbf{R}'', ``\textbf{F}'', ``\textbf``{I}''	\\
			~~~~~followed by three axis lengths and three angles (in degrees)\\
			\underline{\textbf{semicolon}}: 
			lines beginning with a semicolon are treated as comments\\
			\underline{\textbf{END:}} ends the data input section\\
			\bottomrule
		\end{tabular}

	\end{table}	
	\subsection{Examples of unit cell inputs}
	\noindent
	P 10 20 30 90 111 90\\
	G 100 400 900  0 -215.02 0\\
	S6   0 -107.51  0   7.51 -400 -792.49\\
	; this is a comment

	\section{Programs}

	\begin{table}
		\caption{Programs for convert lattice representations}
		\begin{tabular}{l l l p{.2\textwidth} p{.5\textwidth} l}
			\toprule
			Name&	in&	out&	param& Output	\\
			\midrule
			CmdToB4&	y&	\BIV{}&&		For each input, 
			it produces the 4 vectors as a,b,c,d as E3 vectors, 
			and also the lengths of each of those vectors.\\[.9pt]
			CmdToC3&	y&	\CIII{}&&		
			For each input, produces the \CIII{} representation in the form 
			(\#,\#) (\#,\#) (\#,\#), 
			which does not conform to the format for input to other programs.\\[.9pt]
			CmdToCell&	y&	a,b,c,$\alpha$,$\beta$,$\gamma$&&		
			Converts to the conventional unit cell representation.\\[.9pt]
			CmdToD13&	y&	\DXIII{}&&		
			Outputs the lengths of the 13 unique vectors describing the Dirichlet cell.\\[.9pt]
			CmdToDC&	y&	y		\\[.9pt]
			CmdToG6&	y&	\GVI{}&&		Converts the input to \GVI{}\\[.9pt]
			CmdToS6&	y&	\SVI{}&&		Converts the input to \SVI{}\\[.9pt]
			\bottomrule

		\end{tabular}
	\end{table}

		\begin{table}
		\caption{Summary of programs manipulating data}
		\begin{tabular}{l l l p{.2\textwidth} p{.5\textwidth} l}
			\toprule
			Name&	in&	out&	param& Output	\\
			\midrule
			CmdCmplx&	NA&	y&&Currently, this just outputs some 
			programmed examples.		\\[.9pt]
			CmdDelone&	y&	\SVI{}&&Converts input to\SVI{}		\\[.9pt]
			CmdGen&	NA&	\GVI{}&	ngen& - number of examples to generate (optional)\\
			{}&&&type - &lattice type to generate, many options (optional) 
			c,t,h,o,r,h,m,a, 
			cP, cF,cI, tP, tI, hP, hR, oP, oF, oI, oS, mP, mC, mS, aP, 
			numeric Niggli types - 1-44, 
			Delone types - C1,C3,C5,~\dots 
			If no parameters are present, one each of the 44 Niggli types. 
			If only a number is present, that is how many of each of the 44 Niggli types to generate.
			For usage examples, see Table \ref{CMDGENexamples}\\[.9pt]
					CmdLM&	y&	cy&&		Lattice Matching. 
			The first input cell is used as the reference. 
			Succeeding cells are matched 
			as well as possible to the reference cell.\\[.9pt]
			CmdNiggli&	y&	\GVI{}&&		Produces the Niggli reduced cell for each input\\[.9pt]
			CmdPath&	?&	? && Clearly not working the way it should		\\[.9pt]
			CmdPerturb&	y&	y&	n-number of perturbations, parts per thousand to perturb&
			n cells perturbed normal to the input \SVI{} vector
			If a lattice centering (including P) is input,
			the output is in the same centering. Otherwise, the output is in \SVI{}.	\\[.9pt]
			CmdS6Refl&	y&	\SVI{}&	&	For each input, 
			all 24 \SVI{} reflections are produced.\\[.9pt]
			CmdSella&	y&	y		\\[.9pt]
			CmdSort&	y&	y&	"C3" or "seq'&	\\[.9pt]
			CmdVolume&	y&	mod&&		Outputs the unit cell and the volume of each input cell.\\[.9pt]
			MultiMetricDists&	y& mod&"Y/y"(optional)&	Parameter Y/y for maxima style output		\\[.9pt]
			PlotC3&	y&	mod& 
			1,2,or 3 for which \CIII{} coordinate to plot. 
			"seq" or "C3"& Plot one of the 3 \CIII{} coordinates, 
			sorted by minimizing point to point distance 
			for \SVI{}{ reflections or 
				sorting the \CIII{} components to put them 
				in the asymmetric unit.}	\\[.9pt]
			Radial&	y&	as radial	&&	
			Radial computes the polar distances in Angstroms 
			from the first input cell. 
			That is {a,$\alpha$}, {b,$\beta$}, and {c,$\gamma$} 
			as coordinates in complex space.\\[.9pt]
			SELLA&	y&	mod&&	
			Outputs the distances from each of the Bravais lattice types 
			(in \svi{}).	\\[.9pt]
			SVD&	y&	mod&& Outputs the singular value decomposition vectors
			and eigenvalues for the input lattice. This is done in \SVI{}.		\\[.9pt]			
			\bottomrule

		\end{tabular}
	\end{table}

	\begin{itemize}
	\item{CmdGen}

CmdGen is a program for creating examples of various types of lattice types.

There are two optional input parameters. The first is the count
of how many samples of each type to generate. The second (if present) is
the type of lattices to generate.  There is no input data
other than the command line parameters.

For examples of using CmdGen, see table \ref{CMDGENexamples}

The is considerable flexibility in the input types. However, they
are case-sensitive.

\begin{itemize}
	\item Niggli types by number 1-44 (see Table \ref{NiggliFormsI}).
	\item Delone types by Delone's designation - 
	C1,C3,C5, T1, T2, T5, R1, R3,\\ 
	O1A, O2, O3, O4, O5, O1B, M1A, M2A, M3, M4, M1B, M2B, 
	A1, A2, A3, H4 (see Table \ref{s6table}).
	\item Crystal system - c,h,t,o,m,a
\end{itemize}

\item SVD

On input to a number of lattices, 
SVD calculates a Singular Value Decomposition.
The output consists of the eigenvalues and 
the six eigenvectors.
It is important to check that the form of
each cell is the same; for instance, monoclinic
cell should have the unique interaxial angle
in the same place (\textit{e.g.} $\beta$)

As an example, consider using SVD to compute the information
for one of the

\textbf{Command line:}\\
CmdGen 5 33 | CmdSort seq | SVD\\
; SVD\\
S6 from input cell   0.000 -31.196   0.000 -71.963 -109.653 -131.880\\
S6 from input cell   0.000 -23.060   0.000 -77.367 -110.398 -125.599\\
S6 from input cell   0.000 -28.898   0.000 -54.985 -106.336 -82.229\\
S6 from input cell   0.000 -56.133   0.000 -103.736 -177.343 -123.204\\
S6 from input cell   0.000 -10.111   0.000 -50.618 -91.57100 -98.102\\

\textbf{eigenvalue     0}         vector         1 0 0 0 0 0\\
eigenvalue 413.012       vector         0 -0.1728 00 -0.4007 -0.6616 -0.6097\\
eigenvalue 48.3869       vector         0 -0.4009 0 -0.1129 -0.5164 0.7482\\
eigenvalue 14.0785       vector         0 -0.7060 0 -0.4581 0.5344 -0.0786\\
eigenvalue 8.53435       vector         0 0.5576 0 -0.7853 0.1001 0.2494\\
\textbf{eigenvalue     0}         vector         0 0 -1 0 0 0

The two zero eigenvalues correspond to the invariant
interaxial angles ($\alpha$ and $\gamma$, in this case).

\end{itemize}

\vspace{2pt}

\begin{table}

		\label{CMDGENexamples}
		\caption{CmdGen examples}

		\begin{tabular}{l l l l}
			\midrule
			CmdGen & generates a single example of each of the 44 Niggli types\\[2pt]
			CmdGen 2 & generates two examples of each of the 44 Niggli types\\[2pt]
			CmdGen 4 aP & generates four examples of each of the two anorthic Niggli types\\[2pt]
			CmdGen 2 17& generates 2 examples of Niggli 17, which is mC\\
			{}&\textbf{Output}:\\
			{}&; Niggli lattice type requested\\
			{}&; lattice type = 17\\
			{}&G6    90.428    90.428   142.222    83.992    83.992    22.888   IT\# = 17  mC\\
			{}&G6    79.033    79.033   111.127   -61.421   -73.872   -22.772   IT\# = 17  mC\\
			CmdGen 2 C5& generate two examples of Delone type "C5", which is
			primitive cubic.\\
			{}&\textbf{Output}:\\
			{}&; Delone lattice type input\\
			{}&; lattice type = C5\\
			{}&G6   181.275   181.275   181.275     0.000     0.000     0.000   IT\# = C5  cP\\
			{}&G6    74.526    74.526    74.526     0.000     0.000     0.000   IT\# = C5  cP\\[2pt]

			CmdGen 1 h& Generate a single example of each of the hexagonal (and rhombohedral) Bravais lattice (per Niggli).\\
			{}&\textbf{Output}:\\
			{}&; Niggli lattice type input\\
			{}&; lattice type = 2\\
			{}&G6   119.828   119.828   119.828   -45.302   -45.302   -45.302   IT\# = 2  hR\\
			{}&;\\
			{}&; lattice type = 4\\
			{}&G6    86.837    86.837    86.837    19.813    19.813    19.813   IT\# = 4  hR\\
			{}&;\\
			{}&; lattice type = 9\\
			{}&G6     3.688     3.688   145.128     3.688     3.688     3.688   IT\# = 9  hR\\
			{}&;\\
			{}&; lattice type = 12\\
			{}&G6    55.273    55.273   136.854     0.000     0.000   -55.273   IT\# = 12  hP\\
			{}&;\\
			{}&; lattice type = 22\\
			{}&G6    73.578    73.578   149.339     0.000     0.000   -73.578   IT\# = 22  hP\\
			{}&;\\
			{}&; lattice type = 24\\
			{}&G6   101.390   125.771   125.771   -91.974   -67.593   -67.593   IT\# = 24  hR\\
			{}&;\\
			\bottomrule
		\end{tabular}	
	\end{table}

		\begin{table}
		\caption{The table of \citeasnoun{Delaunay1932} describing the 24
			Bravais types in \SVI{}. It has been redone removing the images of the 
			Dirichlet cells, the not-reduced cells, and adding the ''lattice character", which
			describes the linear manifold of each type. The crystal family types have been
			renamed to modern usage: Q changed to T for tetrahedral, K changed
			to C for cubic, and T changed to A for anorthic. Where Delone
			in some places included two types in one table cell, they have been
			split into two (for example: ''M1" becomes ''M1A" and ''M1B").

			Note that five types (O3, M3, M2B, A2, and A3) are not normal 
			crystallographic types. They are boundary types, and they have
			fewer free parameters than the generic type requires. For instance,
			O3 (character: rs0 rs0) has only two free parameters (r and s), whereas
			an ordinary orthorhombic type requires three variables.}
		\includegraphics[width=0.75\textwidth]{DeloneTypesGrid_3}\\
		\label{s6table}		
	\end{table}

\begin{table}
	\caption{Roof/Niggli symbol, International Tables (IT) lattice character, Bravais lattice type, unsorted ${\mathbf{DC}^{\mathbf{7}}}$ subspace, boundary polytope.
		Note that the variables r, s and t are non-negative, and u, v and w may be
		positive, negative or zero as constrained below.}
	\begin{center}
		\begin{tabular}{|c|c|c|c|c|}
			\hline
			{\bf Roof/}     &{\bf IT}&{\bf Bravais}&{\bf Unsorted ${\mathbf{DC}^{\mathbf{7}}}$}&{\bf Bound-}\\
			{\bf Niggli}&{\bf Lattice}&{\bf Lattice}&{\bf Subspace}&{\bf ary}\\
			{\bf Symbol}&{\bf Char}&{\bf Type}&                             & {\bf Polytope} \\
			\hline
			44A&3&${\bf cP}$&$(r,r,r,2r,2r,2r,3r)$&$12345\!=\!12\hat{3}\!=\!12\hat{4}\!=\!12\hat{5}$\\
			\hline
			44C&1&$cF$&$(r,r,r,r,r,r,2r)$&12679ACD\\
			\hline
			44B&5&$cI$&$(r,r,r,4 r/3,4 r/3, 4 r/3,r)$&$\text{12F2}^{\prime} \text{F}^{\prime} = 1\hat{2}\hat{\text{F}}$\\
			\hline
			45A&11&${\bf tP}$&$(r,r,t,r+t,r+t,2r,2r+t)$&$1345 = 1\hat{3} = 1\hat{4} = 1\hat{5}$\\
			45B&21&${\bf tP}$&$(r,s,s,2s,r+s,r+s,r+2s$&$2345 = 2\hat{3} = 2\hat{4} = 2\hat{5}$\\
			\hline
			45D&6&$tI$&$(r,r,r,r-w/2,r-w/2,2r+w,r),$&\\
			&&&$-r \leq w \leq 0$&$\text{12FF}^{\prime} = 12\hat{\text{F}}$\\
			45D&7&$tI$&$[r,r,r,2*r+u,r-u/2,r-u/2,r],$&\\
			&&&$-r \leq u \leq 0$&$\text{12F2}^{\prime} = 1{\hat{2}}\text{F}$\\
			45C&15&$tI$&$(r,r,t,t,t,2r,t)$&158BF\\
			45E&18&$tI$&$(r,s,s,-r/2+2s,s,s,-r/2+2s)$&$\text{2ADA}^{\prime}  = 2{\hat{A}}\text{D}$\\
			\hline
			48A&12&${\bf hP}$&$(r,r,t,r+t,r+t,r,r+t])$&134E\\
			48B&22&${\bf hP}$&$(r,s,s,s,r+s,r+s,r+s)$&2458\\
			\hline
			49C&2&$hR$&$(r,r,r,2r-u,2r-u,2r-u,3r-u),$&\\
			&&&$0 < u \leq r$&$121^\prime 2^\prime = \hat{1}\hat{2}$\\
			49D&4&$hR$&$(r,r,r,2r+u,2r+u,2r+u,3r+3u),$&\\
			&&&$-r \leq u \leq 0$&$121^\prime 2^\prime = \hat{1}\hat{2}$\\
			49B&9&$hR$&$(r,r,t,t,t,r,r+t)$&1679ACD\\
			49E&24&$hR$&$(r,s,s,s+r/3,s+r/3,s+r/3,s)$&$2F2^\prime \text{F}^\prime = \hat{2}\hat{\text{F}}$\\
			\hline
			50C&32&${\bf oP}$&$(r,s,t,s+t,r+t,r+s,r+s+t)$&$345 = \hat{3} = \hat{4} = \hat{5}$\\
			\hline
			50D&13&$oC$&$(r,r,t,r+t,r+t,2r+w,2r+t+w),$&\\
			&&&$-r \leq w \leq 0$&134\\
			50E&23&$oC$&$(r,s,s,u+2*s,s+r,s+r,u+2*s+r),$&\\
			&&&$-s \leq u \leq 0$&245\\
			50A&36&$oC$&$(r,s,t,s+t,t,r+s,s+t)$&35B\\
			50B&38&$oC$&$(r,s,t,s+t,r+t,s,s+t)$&34E\\
			50F&40&$oC$&$[r,s,t,t,r+t,r+s,r+t)$&458\\
			\hline
			51A&16&$oF$&$(r,r,s,r+s+u,r+s+u,-2u,s),$&\\
			&&&$-r \leq u \leq 0$&$\text{1F1}^{\prime} = \hat{1}\text{F}$\\
			51B&26&$oF$&$(r,s,t,-r/2+s+t,t,s,-r/2+s+t)$&$\text{ADA}^{\prime}  = \hat{A}\text{D}$\\
			\hline
			52A&8&$oI$&$(r,r,r,2r+u,2r+v,-u-v,r),$&\\
			&&&$-r \leq u \leq 0, -r \leq v \leq 0$&12F\\
			52B&19&$oI$&$ (r,s,s,2s-u,s,s,-r+2s+u),$&\\
			&&&$0 < u \leq r $&\\
			&&&$ (r,s,s,2s-u,s,s,r+2s-u),$&\\
			&&&$r < u \leq s $&29C = 2AD\\
			52C&42&$oI$&$r,s,t,t,t,r+s,t$&58BF\\
			\hline
		\end{tabular}
	\end{center}
	\label{NiggliFormsI}
\end{table}%

	\section{Availability of code}

	The $C^{++}$ ~code for C++ is available in github.com, in
	\url{https://github.com/duck10/LatticeRepLib.git}.

	\ack{{\bf Acknowledgements}}

	\ack{{\bf Funding information}}

	Funding for this research was provided in part by:  
	US Department of Energy Offices of Biological and 
	Environmental Research and of Basic Energy Sciences 
	(grant No. DE-AC02-98CH10886; grant No. E-SC0012704); 
	U.S. National Institutes of Health (grant No. P41RR012408; 
	grant No. P41GM103473; grant No. P41GM111244; 
	grant No. R01GM117126,
	grant No. 1R21GM129570); Dectris, Ltd.

	%	\bibliography{reduced}
	%	
	%	\bibliographystyle{iucr}

	%-------------------------------------------------------------------------
	% TABLES AND FIGURES SHOULD BE INSERTED AFTER THE MAIN BODY OF THE TEXT
	%-------------------------------------------------------------------------

	% Simple tables should use the tabular environment according to this
	% model

	% Postscript figures can be included with multiple figure blocks

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%