### 2025-04-02 SELLA/Sella.tex
LinearAxis, Claude's attempt to help me write an article. SELLA -- updating the SELLA article. More to do

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2013 International Union of Crystallography
% Version 1.6 (28 March 2013)
%------------------------------------------------------------------------------

\documentclass[preprint]{iucr}              % DO NOT DELETE THIS LINE
%\documentclass[]{iucr}              % DO NOT DELETE THIS LINE
\usepackage{amssymb}
\usepackage{bm}
\usepackage[fleqn]{amsmath}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `basename #1 .tif`.png}
%\usepackage{amsbsy}
\usepackage{float}
\usepackage{url}
%\usepackage{yfonts}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[outdir=./]{epstopdf}
\usepackage{epsfig}
\usepackage{booktabs}

\newcommand{\SVI}[0]{$\mathbf{S^{6}}$}
\newcommand{\GVI}[0]{$\bf{G^{6}}$}
\newcommand{\scalarsub}[2]{$#1_#2$}
\newcommand{\vdotv}[2]{${{\bf #1 \cdot #2}}$}
\newcommand{\charseq}[1]{({\bf #1})}

%-------------------------------------------------------------------------
% Information about journal to which submitted
%-------------------------------------------------------------------------
\journalcode{A}              % Indicate the journal to which submitted
%   A - Acta Crystallographica Section A
%   B - Acta Crystallographica Section B
%   C - Acta Crystallographica Section C
%   D - Acta Crystallographica Section D
%   E - Acta Crystallographica Section E
%   F - Acta Crystallographica Section F
%   J - Journal of Applied Crystallography
%   M - IUCrJ
%   S - Journal of Synchrotron Radiation

\begin{document}                  % DO NOT DELETE THIS LINE

	%-------------------------------------------------------------------------
	% The introductory (header) part of the paper
	%-------------------------------------------------------------------------

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).

	{\LARGE \emph{\today}} \\
	\title{Determining Bravais Lattice Types}
	%\shorttitle{SELLA  Determining Bravais Lattice Types}

	% Authors' names and addresses. Use \cauthor for the main (contact) author.
	% Use \author for all other authors. Use \aff for authors' affiliations.
	% Use lower-case letters in square brackets to link authors to their
	% affiliations; if there is only one affiliation address, remove the [a].

	\cauthor[a]{Lawrence C.}{Andrews}{larry6640995@gmail.com}{}
	\author[b]{Herbert J.}{Bernstein}
	\author[c]{Nicholas K.}{Sauter}

	% ORCID for LCA - 0000-0002-4451-1641

	\aff[a]{9515 NE 137th St, Kirkland, WA, 98034-1820 \country{USA}}
	\aff[b]{c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}, Rochester Institute of Technology, c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}}
	\aff[c]{Lawrence Berkeley National Laboratory, 1 Cyclotron Rd., Berkeley, CA, 94720 \country{USA}}

	% Use \shortauthor to indicate an abbreviated author list for use in
	% running heads (you will need to uncomment it).

	\shortauthor{Andrews, Bernstein, and Sauter}

	% Use \vita if required to give biographical details (for authors of
	% invited review papers only). Uncomment it.

	% lca IUCr id IUCr6401
	%\vita{Author's biography}

	% Keywords (required for Journal of Synchrotron Radiation only)
	% Use the \keyword macro for each word or phrase, e.g. 
	% \keyword{X-ray diffraction}\keyword{muscle}

	%\keyword{keyword}
	\keyword{Unit cell reduction}
	\keyword{Delaunay}
	\keyword{Delone}
	\keyword{Niggli}
	\keyword{Selling}
	\keyword{Bravais Lattice}

	% PDB and NDB reference codes for structures referenced in the article and
	% deposited with the Protein Data Bank and Nucleic Acids Database (Acta
	% Crystallographica Section D). Repeat for each separate structure e.g
	% \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

	%\PDBref[optional name]{refcode}
	%\NDBref[optional name]{refcode}

	\maketitle                        % DO NOT DELETE THIS LINE

	\begin{synopsis}
		A method for determining likely Bravais lattice types based on Selling (Delone) 
		reduction. It is a complete, closed solution.
	\end{synopsis}

	\begin{abstract}
		We introduce a new Bravais lattice determination algorithm. SELLA is a 
		straight-forward algorithm and a program for determining Bravais lattice type 
		based on Selling (Delone) reduction. It provides a clear metric of fit to each type. The mathematical foundations for treating near-symmetries are
		described. The 24 Bravais lattice types described by \citeasnoun{Delaunay1932} 
		are described as linear manifolds in 6-dimensions.
	\end{abstract}

	%-------------------------------------------------------------------------
	% The main body of the paper
	%-------------------------------------------------------------------------
	% Now enter the text of the document in multiple \section's, \subsection's
	% and \subsubsection's as required.

	\section{Introduction}

	We introduce a new Bravais lattice determination algorithm. The Bravais 
	lattice types were created by \citeasnoun{Bravais1850}. \citeasnoun{Delaunay1932} 
	and \citeasnoun{Niggli1928} developed methods for the identification 
	of the Bravais lattice type of a crystal using the measured unit cell 
	dimensions; their methods were exact only if the cell parameters 
	exactly corresponded to the actual type \cite{Patterson1957}. 
	Delaunay \cite{Delaunay1932} discussed the issues of scalars that have nearly zero values, 
	but the broader issues of measurement errors were not 
	discussed. (Delaunay later changed the spelling to Delone, a
	transliteration of the Cyrillic spelling of his name.)

	\citeasnoun{Zimmermann1985} described the program, DELOS
	for the same purpose and based on the use of Selling scalars, which
	SELLA uses as well; they treat the scalars as a list of 6 numbers, but do not consider them as points in 
	a space.  SELLA treats the scalars as elements of vectors in the 6-dimensional
	space, \SVI. Treating lattices as points in a liner space accesses the tools
	and systematic procedures of linear algebra.

	\citeasnoun{andrews2014} reviewed the literature on 
	 efforts to create methods to utilize data that contain unavoidable 
	measurement error and adds a more generic view of the problem.

	SELLA is a based on Selling (Delone) reduction. It is a complete, closed solution, and it provides a clear metric of fit to each type.

	\section{Terminology}

	\subsection{The Bravais types}

	\citeasnoun{Delaunay1932} found that the 14 Bravais lattice types 
	could be allocated among 24 types that are determined by the 
	type of Dirichlet unit cell (termed the Voronoi region \cite{Voronoi1908} or 
	Dirichlet region \cite{dirichlet1850} by 
	Delone). Delone used the convention of \citeasnoun{Bravais1850}
	using the sides of the tetrahedron labeled by the symbols of the six
	Selling scalars. Figure \ref{fig:PQRSTU} shows the labeling used in
	the International Tables for Crystallography \cite{Henry1952}; 
	Table \ref{table:ScalarConcordance} shows other choices that have been used. The
	contents for each type are described in Figure \ref{fig:PQRSTU}.

	\citeasnoun{Burzlaff1985} revised the labeling of the types.
	Here we have returned to the numbering of the Bravais types to that 
	of \citeasnoun{Delaunay1932}.
	In some cases, Delone included two forms within a single 
	type. \citeasnoun{Burzlaff1985}
	renumbered by ordinals. We have labeled the Delone 
	types that have multiple items as "A" and "B". Table \ref{table:DeloneTypeConcordance}
	shows the various labelings. Figure \ref{fig:rotatedDeloneTypes} 
	revises the figure of \citeasnoun{Delaunay1932}, using 
	more modern notation. Table \ref{table:ScalarConcordance} 
	lists the various labelings that have been 
	used for the sides of the Bravais tetrahedron. 
	Figure \ref{fig:PQRSTU} shows the arrangement of the 
	symbols as used in the International Tables for Crystallography.

	\begin{table}
		\begin{tabular}{cccccccccccc} 
			\toprule

			\rotatebox{80}{\citeasnoun{Delaunay1932}}
			& \rotatebox{80}{\citeasnoun{Henry1952}  }
			& \rotatebox{80}{\citeasnoun{Patterson1957} }
			& \rotatebox{80}{\citeasnoun{Burzlaff1985}}
			&\rotatebox{80}{\citeasnoun{andrews2019} } \\	
			\midrule
			g&P&h$_{23}$&$a_2.s_3$&\scalarsub{s}{1}\\		
			k&R&h$_{12}$&$a_1.a_3$&\scalarsub{s}{3}\\		
			l&S&h$_{14}$&$a_1.a_2$&\scalarsub{s}{4}\\		
			m&T&h$_{24}$&$a_1.a_4$&\scalarsub{s}{5}\\		
			h&Q&h$_{13}$&$a_2.a_4$&\scalarsub{s}{2}\\		
			n&U&h$_{34}$&$a_3.a_4$&\scalarsub{s}{6}\\		
			\bottomrule
		\end{tabular}

		\caption{The various terms used to describe the sides of 
			the Bravais tetrahedron. (In the figures in \citeasnoun{Delaunay1932} 
			the tetrahedron has been rotated by -120 degrees)}
		\label{table:ScalarConcordance}
	\end{table}

	\begin{table}
		\begin{tabular}{ccccccccc}
			\toprule
			\cite{Delaunay1932} & \cite{Burzlaff1985}   &this  \\
			&  & paper \\
			\midrule
			K$_I$ & K1 & C1 & cI \\		
			K$_{III}$ & K2 & C3 &cF \\		
			K$_V$ & K3 & C5 &cP \\		
			Q$_I$ & Q1 & T1 &tI \\		
			Q$_{II}$ & Q2 & T2 &tI \\		
			Q$_V$ & Q3 & T5 & tP\\		
			R$_I$ & R1 & H1 & hR \\		
			R$_{III}$ & R2 & H3 & hR \\		
			O$_I$ & O1 & O1A & oF  \\		
			O$_I$ & O2 & O1B &oI  \\		
			O$_{II}$ & O3 & O2 & oI   \\		
			O$_{III}$ & O4 & O3$^*$ & oI   \\		
			O$_{IV}$ & O5 & O4 & oS   \\		
			O$_V$ & O6 & O5 & oP   \\		
			M$_I$ & M1 & M1A & mC  \\		
			M$_I$ & M2 & M1B & mC \\		
			M$_{II_2}$  & M3 & M2A$^*$ & mC \\		
			M$_{II}$ & M4 & M2B$^*$ & mC  \\		
			M$_{III}$ & M5 & M3 & mC  \\		
			M$_{IV}$ & M6 & M4 & mP  \\		
			T$_I$ & T1 & A1 & aP  \\ 		
			T$_{II}$ & T2 & A2$^*$ & aP \\		
			T$_{III}$ & T3 & A3$^*$ & aP \\		
			H$_{IV}$  & H1 & H4 & hP \\ 
			\bottomrule
		\end{tabular}
		~\\ ~\\
		\caption{The 24 Delone types. \cite{Delaunay1932}\\
			$^*$ not a full Bravais type.}
		\label{table:DeloneTypeConcordance}
	\end{table}

	\subsection{The space \SVI{}}
	\cite{andrews2019} recast the Selling parameters to define a metric space, \SVI{}.
	The base vectors, $\bf{a}$, $\bf{b}$, $\bf{c}$ of the unit cell
	are used to define $\bf{d}=-(\bf{a}+\bf{b}+\bf{c})$.
	A point ${s}$ in \SVI{} is defined as:\\
	${s}$= [\scalarsub{s}{1}, \scalarsub{s}{2}, \scalarsub{s}{3},\scalarsub{s}{4}, \scalarsub{s}{5}, \scalarsub{s}{6}] \\
	where \scalarsub{s}{1}=\vdotv{b}{c}, ~\scalarsub{s}{2}=\vdotv{a}{c}, ~\scalarsub{s}{3}=\vdotv{a}{b}, 
	~\scalarsub{s}{4}=\vdotv{a}{d}, ~\scalarsub{s}{5}= \vdotv{b}{d}, and \scalarsub{s}{6}=\vdotv{c}{d}.

	\subsection{Terminology of the Bravais lattice types}
	The Bravais tetrahedron symbol is described in Figure \ref{fig:PQRSTU}.\\
	The different ways that the scalars have been labeled is shown in Table \ref{table:ScalarConcordance} \\
	The table of the Delone types is displayed in Figure \ref{fig:rotatedDeloneTypes}. The table has been transposed from that of Delone. Where Delone listed 2 types in one cell, we have been split them into parts A and B. The crystal family names and abbreviations have been changed to modern usage. \\

	\subsection{Character}
	In \SVI{} for reduced cells, each Bravais type is a single, connected,
	origin-inclusive, Euclidean manifold. As such, all points in a
	particular manifold can be described by their ``character'', 
	a representation of the vectorsns of that type. The table of \citeasnoun{Delaunay1932} included this information 
	encoded in the formalism of Bravais, as a tetrahedron, with
	the sides marked to show which of the 6 Selling parameters were
	zero or which pairs (or multiplets) have equal values.

	As an example, in Figure \ref{fig:O4_terminology}, the tetrahedron shows two
	zeros, in positions P and Q (\scalarsub{s}{1} and \scalarsub{s}{2}). The slashes on the lines
	show that positions S and T (\scalarsub{s}{4} and \scalarsub{s}{5}) are equal. This is 
	encoded in the \textit{character}, which is shown at the bottom of the item as \charseq{00r sst}.

	\section{Algorithms for creation of an exhaustive list of the manifolds of the Bravais types}

	\subsection{Deriving the projectors and perps}

	We will need to create projectors for each of the Bravais
	lattice types. The projectors will be use to generate
	images of probe points in each of the types, and the perps
	will measure the distance from each manifold.

	We will use Selling reduction; Niggli reduction has a fundamental 
	unit that is non-convex and some boundaries of the fundamental 
	unit are partly closed and partly open \cite{andrews2014} leading 
	to considerable complexity. Because Niggli reduction divides the 
	fundamental unit of Niggli reduction into separate all obtuse and all acute regions, 
	the transitions between those are quite complex \cite{andrews2014}. 
	In space \SVI{} \cite{andrews2019b}, the fundamental 
	unit is the simple all-negative orthant of a 6-dimensional 
	Cartesian space. Only two kinds of operations in the space 
	will need to be considered: the 6  transform operations 
	at the boundaries of the orthant (zeros of an axis) and 
	24 reflections (allowed permutations) \cite{andrews2019}

	Twenty-four lattice types were enumerated by \citeasnoun{Delaunay1932}. 
	Three of those are triclinic, and we will ignore them because all crystals 
	can be indexed as triclinic. We 
	shall need all representations of all non-triclinic lattice 
	types within the \SVI{} fundamental unit (all Selling-reduced 
	cells) or those generated beyond the boundaries by boundary transforms. 
	Three forms (M3, M2B, and O3) are reduced dimension boundaries that do
	not correspond to unique Bravais types; any lattice that falls in those
	three is indexed by one of the other types.

	Seven of the 21 Delone types do not have a zero in their \SVI{} 
	representation; the presence of zero indicates that a form is 
	on the boundary of the \SVI{} fundamental unit. Since a 
	point on the boundary transforms to another boundary 
	point \cite{andrews2019b}, we can generate the Virtual 
	Cartesian Point (VCP) \cite{andrews2019b} appropriate to 
	each zero in the \SVI{} vector. Because the reduction
	operations do not commute, in the cases where there is more 
	than one zero in the vector, all orderings of applying 
	reductions to all of the zeros must be generated. (For example, 
	for two zeros, there will be six vectors produced.) Again, 
	there may be many duplicate results.

	Finally, the projectors must be computed for each of the many 
	sample vectors of the 21 types. We will also need the 
	``perps'', the operators that give the normal vector to each 
	such manifold. For a particular manifold, the projector 
	applied to the probe gives
	the least-squares best fit point within that manifold
	to the probe vector. 
	Perps are computed by subtracting the 
	corresponding projector from the unit matrix. The norm of avector
	created by multiplying the probe times the projector
	 is the distance from that manifold.

	\subsubsection{Generating sample vectors from all Bravais types\protect\\}

	Step 1: Generate any random vector that contains no zero 
	values and no duplicate values.  For a later operation, 
	it should only contain values large enough that adding or
	subtracting the smallest
	representable floating point number will not change the 
	value in computer arithmetic ($DBL\_MIN$ in the C language).
	That value will be used to mark the positions of zeros
	in vectors that have been processed.

	Step 2; For each Delone type, multiply its projector times 
	the vector from Step 1. Store  the resulting vector
	from Step 1 in a list 
	with entries for each Bravais type (see Table \ref{table:DeloneTypes} 
	for the projectors).

	Step 3:

	a) If the vector result of Step 2 contains no zeros, generate the 24 reflections
	and store in the list by Bravais type.

	\textit{\textbf{LCA--need to normalize the vectors, just like 
		in lattice matching. ? Can we use the 
		lattice matching code here? This is very 
		incomplete. The processing of multiple zeros is MISSING! But the use of DBL\_MIN is there.}}

	b) If the vector contains one zero, apply the boundary transform 
	operation for that boundary
	to the vector, and
	store the products of the vector and its 24 reflections in the list. Note, 
	this is where the choice of $DBL\_MIN$ has effect; adding or
	subtracting such a small value will not change
	any value except zero.

	c) If the vector contains more than one zero and \underline{not} $DBL\_MIN$, 
	first set one zero to $DBL\_MIN$,  and then return it to 
	step b, using the modified vector. Then set the second zero to $DBL\_MIN$ (with the
	first zero still zero), and send that result to step b. 
	If there is a third zero,  process it the same way as 2 zeros and return to c.
	(Note that no valid unit cell can have more than 3 zeros
	in space \SVI{}.)

	Continue those steps until all the types have been processed. 
	There will be duplicates within each type,
	which will be eliminated later.

	\subsubsection{Generate projectors from sample vectors\\}

	Each vector in each of the lists from the previous steps is a representative of one
	of the Delone types. Although a vector is a one-dimensional manifold,
	these vectors encode the information about the full manifolds of the
	Delone types in their zeros and their duplicate values. Now we 
	discuss the creation of the projector matrices that correspond
	to each vector.

	We begin by defining a function: \\ \relax
	\underline{Fraction} computes the reciprocal of the 
	number of elements that equal \\ \relax
	an input value or zero if there are no duplicates.

	\textit{\textbf{LCA write pseudocode for Fraction.}}

	The following steps will be repeated for each of the vectors in the
	list generated above in order to generate the list
	of the corresponding projector matrices.

	To generate a matrix \textit{m} starting from a vector.

	Step 1: Zero any elements of the vector that have been set to $DBL\_MIN$.

	Step 2: For each of the six scalars

	For i in 1-6
	\textit{m}[i,i] = \{if \textit{Fraction}(s[i]) == 0.0 or abs(s[i]) == 0.0\} then 0.0 else \textit{Fraction}(s[i])

	For i in 1-6
	For j in 1-6
	\textit{m}[i,j] 	= \{if s[i]==s[j] or abs(s[j]) $>$ 0\} then~\textit{Fraction}(s[j]) else 0.0

	Add m to a list of the projectors.

	To finish, the duplicate projectors are removed within each type, 
	leaving 239 non-triclinic projectors. (If the triclinic cases are 
	included, there are 10 more.) See Table \ref{table:TypeCounts}.

	\subsection {Measuring the fit}

	When the library of projectors has been prepared, then
	the distance of any probe from each of the manifolds
	can be computed. 
	The distance from the given manifold is just
	the norm of the perp for that manifold applied to the probe.
	An alternative, scale-free measure is the angle between
	the sample vector and its projected image.

	\section{Degenerate Delone types}

	Three of the 21 non-triclinic Delone types are only boundaries of other types. In Figure 
	\ref{fig:degenerate}, Delone types M2B and M3 are in the row for orthorhombic, and
	O3 is in the row for rhombohedral and tetragonal. These types do not need to
	be searched for in general surveys. If a probe is close to 
	one of these degenerate manifold, it will be close to one
	or both of the adjacent manifolds.

	\section{Verification}

	\subsection{Rationale}

	Most of the proposals for Bravais lattice determination in the 
	presence of experimental error do not have a clear proof that 
	they are complete. For many years there were verbal complaints 
	that available programs would not infrequently miss experimental 
	cases (for example, TRACER \cite{lawton1965} and many anecdotal 
	communications).  To our knowledge, only the method of 
	\cite{OishiTomiyasu2012} has clearly demonstrated completeness
	for exact values.

	The justification that Sella is complete has three parts.

	First: We require that the cell being tested be Selling reduced, and
	all Bravais lattice types within the \SVI{} fundamental unit (all
	\SVI{} scalars negative) to have their projectors generated. Therefore,
	the 24 possible forms (reflections) of a 
	reduced cell will be close to one of the 24 copies of the types. 
	For the reflections, we do not need to consider the case of 
	points near a boundary; the boundary 
	transforms must be treated for that.

	Second: In the case of the boundary transform operations, 
	we take the case of a point that is infinitesimally distant from the manifold of a 
	Bravais lattice type that has one zero in its scalars. Take 
	the point in the manifold that is closest to the experimental 
	point. We transform both by the corresponding boundary transform 
	operation. We have generated a new pair of points in a 
	different location, one of which is in some representation of 
	the same Bravais lattice type. The points are still 
	infinitesimally distant from each other.

	Third: Assume we have a point that is outside the fundamental 
	unit and is near an undiscovered manifold for a Bravais lattice 
	type. If the point is near a boundary defined by a single 
	zero, then we apply a boundary transform to the point and to 
	the Bravais manifold, bringing them both into the fundamental 
	unit. If all of the variations of all of the Bravais 
	lattice types within and on the boundaries of the fundamental unit have 
	been generated, then this putative undiscovered manifold 
	has already been included in the list of possibilities. 
	Similar arguments apply to the cases of two or three zeros.

	If we apply all of the boundary transform operations to the 
	representations of the Bravais lattice types, then 
	there are no places that a point in the fundamental 
	unit will fail to find the closest Bravais lattice 
	(or lattices). As stated above, we found 239 
	non-triclinic possible representations.

	What might occur if some reduced point has an unreduced 
	copy near the all-negative orthant?  Would it be 
	possible for it to be close to one of the copies of one 
	of the 21 Delone types? The reduced point is obviously in 
	the non-positive orthant. Further, because all possible 
	copies of all of the Delone type manifolds have been 
	generated, the reduced point will be near a copy of the 
	manifold it was near when it was unreduced.

	\subsection{Testing against known cells}

	89539 unit cells were extracted from the Protein Data Bank 
	in 2018
	(\citeasnoun{bernstein1977} and \citeasnoun{berman2000}). 
	When evaluated with Sella, all but 57 were found at zero 
	distance from the described crystal class. The 57 were 
	identified as incorrect unit cells for the assigned 
	crystal class; however, the distances from the assigned 
	classes were small because the deviations from the 
	appropriate parameters were modest (see Table \ref{tab:PDBCells}).
	The suspect cases have since been corrected \cite{Burley_personal_communication}.
	\begin{table}

		\begin{tabular}{ccccccc}
			\toprule
			PDB ID & Sp.Gr. & Delone type&distance ({{\AA}}) \\ 
			\midrule
			1NUI&P3121&H4&0.0439\\ 
			2NW8&P3121&H4&0.0401\\ 
			2J1L&P3121&H4&0.0307\\ 
			1VYI&P3121&H4&0.2504\\ 
			3FZ8&P32&H4&0.0575\\ 
			2J5W&P3221&H4&0.0543\\ 
			2VAM&P3221&H4&0.1166\\ 
			1Y8J&P3221&H4&0.0389\\ 
			1F4V&P3221&H4&0.0276\\ 
			2ZP9&P6&H4&0.0528\\ 
			1DDR&P61&H4&0.0726\\ 
			1DDS&P61&H4&0.0726\\ 
			1SGU&P61&H4&0.0295\\ 
			3CDC&P61&H4&0.0559\\ 
			2WAG&P6122&H4&0.0377\\ 
			1LBM&P6122&H4&0.0312\\ 
			2O9Z&P62&H4&0.0402\\ 
			1ELZ&P6322&H4&0.4035\\ 
			1SA0&P65&H4&0.0681\\ 
			3KQI&P6522&H4&0.0331\\ 
			2F9Y&P6522&H4&0.0470\\ 
			3E73&P6522&H4&0.0523\\ 
			3CAP&H3&H4&0.0586\\ 
			2UX6&H32&H4&0.0386\\ 
			1ODT&H32&H4&0.0816\\ 
			2FDI&P43&T5&0.0758\\ 
			3HQ7&P43212&T5&0.2869\\ 
			1E2Q&P43212&T5&0.5482\\ \\
			\bottomrule
		\end{tabular}
		\begin{tabular}{ccccccc}
			\toprule
			PDB ID & Sp.Gr. & Delone type&distance ({{\AA}}) \\ 
			\midrule
			1SHN&P43212&T5&0.0491\\ 
			1C50&P43212&T5&1.9040\\ 
			2A1L&P43212&T5&0.0528\\ 
			1S2L&P43212&T5&0.0334\\ 
			1H9S&P41&T5&1.0149\\ 
			1H9R&P41&T5&1.0141\\ 
			2F0Q&P41&T5&0.0260\\ 
			3EYM&P41&T5&0.0476\\ 
			1WV5&P41212&T5&0.0250\\ 
			2DF3&P41212&T5&0.0387\\ 
			1L6O&P41212&T5&0.0346\\ 
			3A58&P41212&T5&0.2976\\ 
			2IU9&P41212&T5&0.0373\\ 
			4BTP&P41212&T5&0.0666\\ 
			1W54&P4212&T5&0.0420\\ 
			2PBE&P42212&T5&0.0548\\ 
			3BR5&P42212&T5&0.2203\\ 
			1GMD&P42212&T5&0.1400\\ 
			1GMC&P42212&T5&0.1402\\ 
			2CCN&P42212&T5&0.0447\\ 
			1CE1&P212121&O5&0.6472\\ 
			3DQ2&P212121&O5&0.1058\\ 
			1Z2A&P212121&O5&0.0670\\ 
			2HA3&P212121&O5&0.3070\\ 
			3BNJ&I4122&T2&0.9518\\ 
			4MGP&I-42d&T1&0.2128\\ 
			4ASL&C2221&O4&0.3383\\ 
			3K7M&P432&C5&0.0517\\ 
			3GBN&I213&C1&1.2661\\ 
			\bottomrule
		\end{tabular}

		\caption{57 unit cells from the PDB around 2018 that did not match their types (later corrected)}
		\label{tab:PDBCells}
	\end{table}
	\subsection{Testing for continuity}

	The manifolds of the Bravais lattice types are characterized by required zeros
	of scalars and/or multiplets of equal values. For instance, all primitive 
	monoclinic cells have two 90 degree angles and so require two zero scalars. Consider H1 as
	an example for multiplets. The character for H1 is \charseq{rrr sss}; two triplets of equal values.

	Consider the case of M4, primitive monoclinic, the character is \charseq{00r stu}. 
	For particular values for r,s,t,u, the best monoclinic cell will not
	depend on whatever values are substituted for the zeros. (It does not depend on the zero values because the projectors to that 
	manifold create zero elements for any vector.) Imangine a plane
	with axis that are the values of the two elements of the character that are zero. Points in the plane represent unit 
	cells with angles defined by the \scalarsub{s}{1} and \scalarsub{s}{2} and r,s,t,u. We can draw a circle in the
	\scalarsub{s}{1},\scalarsub{s}{2} plane, each point
	on the circle corresponding to a lattice.

	Reducing the lattices corresponding to the points on the
	circle and r,s,t,u and calculating
	the distance of the reduced cell from \charseq{00r stu}.
	Plotting the curve of the circle with the distance of
	each point from the origin changed to distance above
	 will give a curve that
	should not have any discontinuities. See Figure \ref{fig:M4}.

	Similarly, Bravais type O5 has a triplet of zeros, character \charseq{000 rst}. Here we
	can generate a sphere that surrounds the orthorhombic point. See Figure \ref{fig:O5}.

	Bravais type O4 (side-centered orthorhombic) has two zeros and a pair of required equal values, 
character \charseq{00r~sst}. The zeros can be treated the same as above
for the case of M4 (primitive monoclinic). 
	The pair of equal values must 
	have an average value that remains equal to $\bf{s}$ in the lattice character. 
	That means 3 values to vary (the 2 zeros and the distance fomaverage
	value for the equal scalars), and we can create a sphere around 
	the point \charseq{00r~sst}. See Figure \ref{fig:O4}.

	\section{Comparisons}
	\textit{\textbf{LCA this section is incomplete. Table needs caption/label.}}
	Comparison to other distance measure; Data for other measures taken
	from \citeasnoun{andrews2014}.

\begin{table}
	\begin{tabular}{cccccc}
		\toprule
		Lattice character & \GVI{} distance & BGAOL Z score & XDS QI & Scaled QI & Sella \\
		\midrule
		mP               &     20.138  & 0.657    &   1.0  & 0.13  & 1.06\\	
		oC               &     125.958 &  3.560   &    23. & 8 3.09& 13.82\\	
		mC               &     125.150 &  4.085   &    23. & 4 3.03& 13.80\\	
		\bottomrule
	\end{tabular}
	\caption{Comparison of distance measures}
	\label{tab:comparisons}
\end{table}
	~\\

	\section{Interpreting the results}

	\section{What could go wrong?}

	% Appendices appear after the main body of the text. They are prefixed by
	% a single \appendix declaration, and are then structured just like the
	% body text.

	%-------------------------------------------------------------------------
	% The back matter of the paper - acknowledgements and references
	%-------------------------------------------------------------------------

	% Acknowledgements come after the appendices

	\ack{{\bf Acknowledgements}}
	Careful copy-editing and corrections by Frances C. Bernstein are 
	gratefully acknowledged.  	Our thanks to Jean Jakoncic and Alexei Soares for 
	helpful conversations and access to data and facilties at 
	Brookhaven National Laboratory.
	% References are at the end of the document, between \begin{references}
		% and \end{references} tags. Each reference is in a \reference entry.

	\ack{{\bf Funding information}}

	Funding for this research was provided in part by: US Department of Energy Offices of Biological and Environmental Research and of Basic Energy Sciences (grant No. DE-AC02-98CH10886 ; grant No. E-SC0012704); U.S. National Institutes of Health (grant No. P41RR012408; grant No. P41GM103473; grant No. P41GM111244; grant No. R01GM117126, grant No. 1R21GM129570; grant No. P30GM133893); Dectris, Ltd.

	\bibliographystyle{iucr}
	\bibliography{Reduced}

	%-------------------------------------------------------------------------
	% TABLES AND FIGURES SHOULD BE INSERTED AFTER THE MAIN BODY OF THE TEXT
	%-------------------------------------------------------------------------

	% Simple tables should use the tabular environment according to this
	% model

	\begin{table}
		\caption{The 21 non-triclinic Delone types with the count of representations.
			The asterisks indicate non-crystallographic types.}
		\label{table:TypeCounts}
		\begin{tabular}{lr}
			Delone type & Count    \\
			\midrule
			H4&12\\
			C1&1\\
			C3&3\\
			C5&16\\
			H1&4\\
			H3&12\\
			T1&3\\
			T2&6\\
			T5&48\\
			O1A&3\\
			O1B&1\\
			O2&6\\
			O3*&9\\
			O4&36\\
			O5&16\\
			M1A&6\\
			M1B&3\\
			M2A&12\\
			M2B*&12\\
			M3*&18\\
			M4&12\\
			\bottomrule
		\end{tabular}
	\end{table}

	\fontseries{Courier}{
		\begin{table}
			\caption{Projectors for the 24 Delone types}
			\label{table:DeloneTypes}
			\begin{tabular}{lcl}
				\toprule
				C1  & (cI) & [1 1 1 1 1 1/  1 1 1 1 1 1/  1 1 1 1 1 1/  1 1 1 1 1 1/  1 1 1 1 1 1/  1 1 1 1 1 1] \\ 
				C3  & (cF) & [1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 0 0 0 0/  1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 0 0 0 0] \\ 
				C5  & (cP) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 1 1 1/  0 0 0 1 1 1/  0 0 0 1 1 1] \\ \\

				T1  & (tI) & [1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 1 0 0 1/  1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 1 0 0 1] \\ 
				T2  & (tI) & [1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 0 0 0 0/  1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 0 0 0 1] \\ 
				T5  & (tP) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 1 1 0/  0 0 0 1 1 0/  0 0 0 0 0 1] \\ \\

				H1  & (hR) & [1 1 1 0 0 0/  1 1 1 0 0 0/  1 1 1 0 0 0/  0 0 0 1 1 1/  0 0 0 1 1 1/  0 0 0 1 1 1] \\ 
				H3  & (hR) & [1 1 0 0 1 0/  1 1 0 0 1 0/  0 0 0 0 0 0/  0 0 0 1 0 0/  1 1 0 0 1 0/  0 0 0 0 0 0] \\
				H4  & (hP) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 1 1 1 0/  0 0 1 1 1 0/  0 0 1 1 1 0/  0 0 0 0 0 1] \\ \\

				O1A & (oF) & [1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 1 0 0 0/  1 1 0 1 1 0/  1 1 0 1 1 0/  0 0 0 0 0 1] \\ 
				O1B & (oI) & [1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 1 0 0 1/  1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 1 0 0 1] \\ 
				O2  & (oI) & [1 0 0 0 1 0/  0 1 0 1 0 0/  0 0 0 0 0 0/  0 1 0 1 0 0/  1 0 0 0 1 0/  0 0 0 0 0 1] \\ 
				O3  & (oI) & [1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 0/  1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 0] \\ 
				O4  & (oS) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 1 0 0 0/  0 0 0 1 1 0/  0 0 0 1 1 0/  0 0 0 0 0 1] \\ 
				O5  & (oP) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 1] \\ \\

				M1A & (mC) & [1 1 0 0 0 0/  1 1 0 0 0 0/  0 0 1 0 0 0/  0 0 0 1 1 0/  0 0 0 1 1 0/  0 0 0 0 0 1] \\ 
				M1B & (mC) & [1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 1 0 0 0/  1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 1] \\ 
				M2A & (mC) & [1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 0/  1 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 1] \\ 
				M2B & (mC) & [1 0 0 0 0 0/  0 1 0 1 0 0/  0 0 0 0 0 0/  0 1 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 1] \\ 
				M3  & (mC) & [1 0 0 0 0 0/  0 1 0 0 1 0/  0 0 0 0 0 0/  0 0 0 1 0 0/  0 1 0 0 1 0/  0 0 0 0 0 0] \\ 
				M4  & (mP) & [0 0 0 0 0 0/  0 0 0 0 0 0/  0 0 1 0 0 0/  0 0 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 1] \\ \\

				A1  & (aP) & [1 0 0 0 0 0/  0 1 0 0 0 0/  0 0 1 0 0 0/  0 0 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 1] \\ 
				A2  & (aP) & [1 0 0 0 0 0/  0 1 0 0 0 0/  0 0 0 0 0 0/  0 0 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 1] \\ 
				A3  & (aP) & [1 0 0 0 0 0/  0 1 0 0 0 0/  0 0 0 0 0 0/  0 0 0 1 0 0/  0 0 0 0 1 0/  0 0 0 0 0 0] \\ \\				
				\bottomrule
			\end{tabular}
		\end{table}
	}
	% Postscript figures can be included with multiple figure blocks

	\begin{figure}
		\caption{Bravais tetrahedron with labeling as used in the International Tables}
		\includegraphics[width=3cm]{PQRSTU}
		\label{fig:PQRSTU}
	\end{figure}

	\begin{figure}
		\caption{The O4 item from the Delone Grid with Delone type description}
		\label{fig:O4_terminology}
		\includegraphics[width=12cm]{O4_terminology}
	\end{figure}

	%	\begin{figure}
		%		\caption{The Delone types}
		%		\label{fig:DeloneTypes}
		%		\includegraphics[width=12cm]{DeloneGridClaude_2-3}
		%	\end{figure}
	\begin{figure}
		\centering
		\includegraphics[angle=90, width=0.7\textheight]{DeloneGridClaude_2-3}
		\caption{The Delone types}
		\label{fig:rotatedDeloneTypes}
	\end{figure}

	\begin{figure}
		\caption{}
		\label{fig:degenerate}
		\includegraphics[width=12cm]{NullSpaceDistribution}
	\end{figure}

	\begin{figure}
		\label{fig:M4}
		\caption{Delone type M4: there are two zeros. A circle of 100 points with
			radius 0.1 is centered at 0,0. The lattices are reduced and the points are
			plotted at the distance from M4.} The red points indicate the original
		circle. The blue points are for the reduced cells.
		\includegraphics[width=12cm]{M4-100-2-zeros}
	\end{figure}

	\begin{figure}
		\caption{Delone type O5: there are three zeros. A sphere 
			is centered at 0,0,0, with radius 0.1. The lattices are 
			reduced, and the point is 
			placed at a distance equal to the distance from type O5. 50,000
			points are plotted.}
		\label{fig:O5}
		\includegraphics[width=12cm]{O5-50K-3-zeros}
	\end{figure}

	\begin{figure}
		\caption{Delone type O4: There are two zeros and one pair.
			A sphere with radius 0.1 was placed around one point. At each
			sphere point, the lattice was reduced and the point on the sphere
			was assigned a color (0.01) to blue (0.0075). 100,000 points are plotted.}
		\label{fig:O4}
		\includegraphics[width=12cm]{O4_Cropped-2-zeros-1-pair}
	\end{figure}

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%